### 2016-05-29 slides.tex

\documentclass{beamer}

\usetheme[numbering=fraction]{metropolis}           % Use metropolis theme
\definecolor{Blue}{HTML}{306D9F}
\definecolor{Orange}{HTML}{CF4A30}

% Theme colors are derived from these two elements
\setbeamercolor{alerted text}{fg=Orange}

\setbeamercolor{normal text}{fg=Blue}
% ... however you can of course override styles of all elements
\setbeamercolor{frametitle}{bg=Blue}

\title{
    \centering
    {\includegraphics[width=0.75\textwidth]{res/logo}}
}

\date{
    \today
}

\author{
    Alberto Amigo Alonso \\
    Sergio Delgado Álvarez \\
    Sergio García Prado \\
    Oscar Fernández Angulo \\
    Silvia Rodriguez Ares \\
}

\institute{Universidad de Valladolid}

\begin{document}
    \maketitle

    \section{First Section}

        \begin{frame}{First Frame}
            Hello, world!
        \end{frame}

    \section{Second Section}

        \begin{frame}{First Frame}
            Hello, world!
        \end{frame}

    \section{Third Section}

        \begin{frame}{First Frame}
            Hello, world!
        \end{frame}

    \begin{frame}[standout]
        ¿Preguntas?
    \end{frame}

\end{document}